---
title: "HW_3_Almond_Profit - Building on HW_2_Almond_Yield"
format: 
  html:
    toc: true
    embed-resources: true
editor: visual
---

Assignment Description:

**1.**  Develop a profit model for your almond yield

Use your imagination to think about what would make sense Make sure you include parameters You could assume a baseline profit and then adjust according to the anomaly

**2.**  Apply the profit model to your almond yield model

There are many ways to combine the almond yield and profit functions; you can have the profit function "call"/use the almond yield function; or create a wrapper function that calls them in sequence (first the almond yield and then the profit function) 

**3.** Perform a simple informal sensitivity analysis of almond yield profit using at least 2 parameters

**4.**  Create 2 graphs
    -   one that shows yield anomaly for each year, accounting for uncertainty in the parameters
    -   one that show how yield anomaly in general varies with your parameters

**5.**  Write a short paragraph (in a Quatro document) to summarize you interpretation of your model results (e.g what do 'take away' from your model and sensitivity analysis)

Submit as a group: a Quatro document, your R files for almond yield and profit model on Canvas - Due Date Monday April 21.

```{r, echo=TRUE, output=FALSE}
# clear the environment
rm(list = ls())
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r, echo=TRUE, output=FALSE}
library(tidyverse)

#library(purrr)

library(here)

```

This assignment aims to model almond almond profit as a function of yield anomalies modeled in the last homework. The almond anomaly model used the regression equation below: 

$$
Y = 0.015T{n,2} + 0.0046T^2{n,2} + 0.07P{1} + 0.0043P^2{1} + 0.28
$$ 

- Y = yield in tons/acre

# Almond Yield Anomaly and Profit Function

```{r, echo=TRUE, output=FALSE}
# load the almond yield and profit wrapper function
source(here("Scripts/Almond_Profit_Wrapper.R"))
```

# Sensitivity Analysis

For our sensitivity analysis, we chose to look at the profit model's sensitivity to the following parameters:

- **Precipitation**: By allowing variation in monthly precipitation in January across our study period, we can assess its impact on both almond yield and profit. We looked at the mean monthly precipitation values during January across each year (1988-2010) in the climate data (`clim.txt`). After analyzing this data, we decided to model uncertainty with this parameter by randomizing with a normal distribution of an 80 mm average with a standard deviation of 40 mm. 

- **Minimum Temperature**: We also looked at the mean monthly minimum temperature values during February across each year (1988-2010) in the climate data (`clim.txt`). After analyzing this data, we decided to model uncertainty with this parameter by randomizing with a uniform distribution across a 7-12 degree Celsius range. 

We generated a total of 2300 samples, which is 100 iterations of the model for each year (23 years total). We used the `pmap` function from the `purrr` package to apply our almond yield and profit function to each of these randomized parameter values. 

```{r, echo=TRUE, output=FALSE}
## Randomization of climate data to obtain uncertainty

# Set seed to get same random values every time
set.seed(80)

#2300 samples; 23 yield anomaly observations (1 each year) times 100 iterations
n_samples <- 2300
year <- rep(1988:2010, 100)

# Randomize for precipitation
precip <- rnorm(n = n_samples, mean = 80, sd = 40)

# Randomize for minimum temperature
min_temp <- runif(n = n_samples, min = 7, max = 12)

# Randomized data with uncertainty parameters
parms <- cbind.data.frame(year, precip, min_temp)

# Randomized data mutated using the almond yield and anomaly function and pmap
results <- parms %>% pmap(almond_yield_anomaly_and_profit)

#Transform results into dataframe containing yield, profit and year
mean_elect <- map_df(results, `[`, c('yield', 'profit'))

# and we can add the parameter values for each run
mean_elect <- cbind.data.frame(mean_elect, parms)

# Check the structure of the data
head(mean_elect)
```

# Plot the Data

```{r}
# Plotting the yield anomaly
Almond_yield_by_year <- ggplot(mean_elect, aes(x = factor(year), y = yield)) +
  geom_boxplot(fill = "lightblue") + #include outliers
  labs(
    title = "Almond Yield by Year",
    x = "Year",
    y = "Yield (tons/acre)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

Almond_yield_by_year
```

```{r}
# Plotting the profit
Almond_profit_by_year <- ggplot(mean_elect, aes(x = factor(year), y = profit)) +
  geom_boxplot(fill = "salmon") + #include outliers
  labs(
    title = "Almond Profit by Year",
    x = "Year",
    y = "Profit ($/acre)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

Almond_profit_by_year
```

# Interpretation of model results

The profit model is based on the yield model, which is a function of temperature and precipitation. The profit model assumes a baseline yield of 0.9 tons per acre, a price of $1000 per ton, and a profit margin of 50%. The profit model is relatively sensitive to changes in these parameters, as shown by the box plots. 

The horizontal lines each year show the median, the boxes show the 50th to 75th percentiles, the whiskers show the 25th and 100th percentiles, and the points are outliers. Our box plots show relatively consistent yields and profits per acre across the time series (1988-2010), with some uncertainty and relatively few outliers. Longer boxes indicate greater uncertainty as the median points fall within a wider range. Same with longer whiskers. Median values and box size stay fairly consistent throughout the time series, indicating low sensitivity to different model inputs.






