---
title: "HW_3_Almond_Profit - Building on HW_2_Almond_Yield"
format: html
editor: visual
---

Assignment:

1.  Develop a profit model for your almond yield

Use your imagination to think about what would make sense Make sure you include parameters You could assume a baseline profit and then adjust according to the anomaly

2.  Apply the profit model to your almond yield model

There are many ways to combine the almond yield and profit functions; you can have the profit function "call"/use the almond yield function; or create a wrapper function that calls them in sequence (first the almond yield and then the profit function) 3. Perform a simple informal sensitivity analysis of almond yield profit using at least 2 parameters

4.  Create 2 graphs
    -   one that shows yield anomaly for each year, accounting for uncertainty in the parameters
    -   one that show how yield anomaly in general varies with your parameters
5.  Write a short paragraph (in a Quatro document) to summarize you interpretation of your model results (e.g what do 'take away' from your model and sensitivity analysis)

Submit as a group: a Quatro document, your R files for almond yield and profit model on Canvas- Due Date Monday April 21

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries

```{r}
library(tidyverse)

#library(purrr)

library(here)

```

This assignment aims to model almond almond profit as a function of yield anomalies modeled in the last homework. The almond anomaly model used the regression equation below: $$
Y = 0.015T{n,2} + 0.0046T^2{n,2} + 0.07P{1} + 0.0043P^2{1} + 0.28
$$ - Y = yield in tons/acre

# Connect Function and Read in Data

```{r almond yield}

#function source
source(here("Scripts/almond_yield.R"))

#read in data

climate_df <- read.table(here("Data/clim.txt"), header = TRUE, sep = "")

# FOR LOOP

# create data frame to store the almond yield results for 1988-2010
yield_results <- data.frame(year = numeric(), yield = numeric())

# create vector of the years 1988-2010
years <- 1988:2010

#run the for loop for each year (yr) of the years vector

for (yr in years) {
  
  #section the data by year
  climate_year <- climate_df[climate_df$year == yr, ]
  
  #run the function using the sectioned data
  yield <- almond_yield(yr, climate_year)  
  
  #store the results in a dataframe
  yield_results <- rbind(yield_results,
                         data.frame(year = yr, yield = yield))
}

print(yield_results)

mean_yield <- mean(yield_results$yield, na.rm = TRUE)

```

# Profit model

```{r almond profit model}

source(here("Scripts/Almond_Profit.R"))

# create data frame to store the almond profit results for 1988-2010
profit_results <- data.frame(year = numeric(), profit = numeric())

# create vector of the years 1988-2010
years <- 1988:2010

#run the for loop for each year (yr) of the years vector

for (yr in years) {
  
  #run the function using the sectioned data
  profit <- profit_from_yield(
    baseline_yield = 1,
    price = 1000,
    yield = yield_results$yield,
    profit_margin = 0.5
  )
  
  #store the results in a dataframe
  profit_results <- rbind(profit_results,
                         data.frame(year = yr, profit = profit))
}

```

# Almond Yield Anomaly and Profit Function
```{r}

almond_yield_anomaly_and_profit <- function(year, precip, min_temp){
  
  coefficients <- c(-0.015, -0.0046, -0.07, 0.0043, 0.28)
  
  yield_anomaly <- coefficients[1] * min_temp + coefficients[2] * min_temp^2 + coefficients[3] * precip + coefficients[4] * precip^2 + coefficients[5]
  
  baseline_yield <- 0.9
  price <- 1000
  profit_margin <- .5
  yield <- (baseline_yield + yield_anomaly)
  
  profit <- (baseline_yield + yield_anomaly) * price * profit_margin 
  return(data.frame(year, yield_anomaly, yield, profit))
}

```

# Sensitivity Analysis

```{r}
set.seed(80)

n_samples <- 2300
year <- rep(1988:2010, 100)

# Randomness for precipitation
precip <- rnorm(n = n_samples, mean = 80, sd = 40)

# Randomness for minimum temperature
min_temp <- runif(n = n_samples, min = 7, max = 12)

parms <- cbind.data.frame(year, precip, min_temp)

results <- parms %>% pmap(almond_yield_anomaly_and_profit)

mean_elect <- map_df(results, `[`, c('yield', 'profit'))
# and we can add the parameter values for each run
mean_elect <- cbind.data.frame(mean_elect, parms)

```




